---
title: "R Notebook"
output: html_notebook
---

car make -- average price
```{r}
library(readr)
library(dplyr)
library(ggplot2)
tc <- read_csv("new_tc.csv")
ordered_grouped_make <- tc %>%
  group_by(Make) %>%
  summarise(average = mean(Price)) %>%
  arrange(desc(average))
write.csv(ordered_grouped_make, file = "ordered_grouped_make.csv")
## most popular brand
ordered <- read_csv("ordered_grouped_make.csv")
most_popular <- tc %>%     ## this will return Ford, which is the most popular car brand
  group_by(Make) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

# part 2
bootstrap correlation test -- test assumption to use regression
```{r}
# select only price, year and mileage
data1 <- tc %>%
  select(Price, Year, Mileage)
n = nrow(data1)
# test between year and price
#w = sample(n,n,replace=T)
#data1_BT = data1[w,]
#cor(data1_BT)[1,2]  # cor between price and year
#cor(data1_BT)[1,3] # cor between price and mileage
#cor(data1_BT)[2,3] # cor between year and mileage

B = 100
data1_cor_BT = matrix(NA, nrow=B, ncol=3) # to compute two correlation

for(i_BT in 1:B){
  w = sample(n,n,replace=T)
  data1_BT = data1[w,]
  data1_cor_BT[i_BT, 1] = cor(data1_BT)[1,2] ## price vs year
  data1_cor_BT[i_BT, 2] = cor(data1_BT)[1,3]  ## price vs mileage
  data1_cor_BT[i_BT, 3] = cor(data1_BT)[2,3]  ## year vs mileage
}
colnames(data1_cor_BT) = c("price vs year", "price vs mileage", "year vs mileage")
head(data1_cor_BT)
quantile(data1_cor_BT[, "price vs year"], c(0.05,0.950)) ## CI for correlation between price and year
quantile(data1_cor_BT[, "price vs mileage"], c(0.05,0.950)) ## CI for correlation between price and mileage
quantile(data1_cor_BT[, "year vs mileage"], c(0.05,0.950)) ## CI for correlation between year and mileage
                                                            ## |abs| less than 0.8 -- no multicollinearity

cor.test(tc$Price, tc$Year,
         alternative = "two.sided",
         method = "pearson")

cor.test(tc$Price, tc$Mileage,
         alternative = "two.sided",
         method = "pearson")

cor.test(tc$Year, tc$Mileage,                 
         alternative = "two.sided",
         method = "pearson")
# bootstrap result:
#       5%       95% 
#0.4039420 0.4088868 
#        5%        95% 
#-0.4246398 -0.1296231 
#        5%        95% 
#-0.7677143 -0.2485699 


```

# display linearity -- yes, linear -- we can fit lm ?? not so sure
```{r}
ggplot(tc, aes(x = Year, y = Price)) + 
  geom_point() + 
  facet_wrap(~Category) 
plot(x = tc$Year, y = tc$Price)
plot(x = tc$Mileage, y = tc$Price)
```

# explore residual plot to see whether wild/residual bootstrap  -- increasing pattern between residual and year, decreasing pattern between residual and mileage -- so we use wild regression
```{r}
n = nrow(tc)
xbar = mean(tc$Price)
price = tc$Price
sd = sd(tc$Price)
res = (price - xbar)/(sd*sqrt(1-1/n))
plot(tc$Year, res)
plot(tc$Mileage, res)
```


# bootstrap regression for low end
```{r}
cate = "low-end"
temp = tc %>% 
  filter(Category == cate) %>% 
  select(Price, Year, Mileage)
fit = lm(Price~Year + Mileage, data=temp)
n1= nrow(temp)
y_predict = predict(fit)
B = 1000
coeff_BT_wild = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict+rnorm(n1, sd=abs(fit$residuals))
  data1_BT = data.frame(years=temp$Year, mileage = temp$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild[,'ce mileage'],c(0.05,0.95))
plot(X,Y, pch=20, cex=0.5, col="gray")
lines(gr_x,gr_y, lwd=2, col="black")
lines(gr_x,gr_y+qnorm(0.999)*gr_y_sd_wild, lwd=2, col="red", lty=2)
#100 bt
#     5%      95% 
#-1413018 -1327120 
#     5%      95% 
#667.8666 710.4538 
#        5%         95% 
#-0.04815179 -0.04355216 
#1000 bt
#      5%      95% 
#-1427509 -1323710 
#      5%      95% 
#666.1856 717.6106 
#         5%         95% 
#-0.04815183 -0.04283102 
```
# bootstrap regression for mid range
```{r}
cate2 = "mid-range"
temp2 = tc %>% 
  filter(Category == cate2) %>% 
  select(Price, Year, Mileage)
fit2 = lm(Price~Year + Mileage, data=temp2)
n2= nrow(temp2)
y_predict2 = predict(fit2)
B = 1000
coeff_BT_wild2 = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict2+rnorm(n2, sd=abs(fit2$residuals))
  data1_BT = data.frame(years=temp2$Year, mileage = temp2$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild2[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild2) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild2[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild2[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild2[,'ce mileage'],c(0.05,0.95))
#      5%      95% 
#-2537737 -2402189 
#      5%      95% 
#1207.071 1274.225 
#         5%         95% 
#-0.06585668 -0.05922462 

```
# bootstrap regression for high end 
```{r}
cate3 = "high-end"
temp3 = tc %>% 
  filter(Category == cate3) %>% 
  select(Price, Year, Mileage)
fit3 = lm(Price~Year + Mileage, data=temp3)
n3= nrow(temp3)
y_predict3 = predict(fit3)
B = 1000
coeff_BT_wild3 = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict3+rnorm(n3, sd=abs(fit3$residuals))
  data1_BT = data.frame(years=temp3$Year, mileage = temp3$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild3[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild3) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild3[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce mileage'],c(0.05,0.95))
#        5%       95% 
# -30520559 -23186964 
#      5%      95% 
# 11628.99 15268.70 
#        5%        95% 
#-1.5287048 -0.3142366      
```


# discard! brand: Ford, training 7/3

```{r}
tesla = tc %>% filter(Make == "Audi")
train_size = nrow(tesla) * 0.7  ## 246
set.seed(100)
trainingRows <- sample(1:nrow(tesla), 246)
train <- tesla[trainingRows, ]
test <- tesla[-trainingRows, ]
```

# discard! convert Mileage and Year to categorical 
```{r}
train$Mileage = factor(train$Mileage)
train$Year = factor(train$Year)
fit = lm(Price ~ Year + Mileage + City, data = train)
summary(fit)
```
# discard! convert make into factor -- one liner regression
```{r}
train_size = nrow(tc) * 0.7  ## 863129
set.seed(100)
trainingRows <- sample(1:nrow(tc), 863129)
train <- tc[trainingRows, ]
test <- tc[-trainingRows, ]

train$Make = factor(train$Make)
fit = lm(Price ~ Year + Mileage + Make, data = train)
summary(fit)
```



