---
title: "R Notebook"
output: html_notebook
---

car make -- average price
```{r}
library(readr)
library(dplyr)
library(ggplot2)
tc_origin <- read_csv("tc.csv")
nrow(tc_origin)
test = tc_origin %>% filter(Make == 'Ferrari')
ordered_grouped_make <- tc %>%
  group_by(Make) %>%
  summarise(average = mean(Price)) %>%
  arrange(desc(average))
write.csv(ordered_grouped_make, file = "ordered_grouped_make.csv")
## most popular brand
ordered <- read_csv("ordered_grouped_make.csv")
most_popular <- tc %>%     ## this will return Ford, which is the most popular car brand
  group_by(Make) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

# part 2
bootstrap correlation test -- test assumption to use regression
```{r}
# select only price, year and mileage
data1 <- tc %>%
  select(Price, Year, Mileage)
n = nrow(data1)
# test between year and price
#w = sample(n,n,replace=T)
#data1_BT = data1[w,]
#cor(data1_BT)[1,2]  # cor between price and year
#cor(data1_BT)[1,3] # cor between price and mileage
#cor(data1_BT)[2,3] # cor between year and mileage

B = 100
data1_cor_BT = matrix(NA, nrow=B, ncol=3) # to compute two correlation

for(i_BT in 1:B){
  w = sample(n,n,replace=T)
  data1_BT = data1[w,]
  data1_cor_BT[i_BT, 1] = cor(data1_BT)[1,2] ## price vs year
  data1_cor_BT[i_BT, 2] = cor(data1_BT)[1,3]  ## price vs mileage
  data1_cor_BT[i_BT, 3] = cor(data1_BT)[2,3]  ## year vs mileage
}
colnames(data1_cor_BT) = c("price vs year", "price vs mileage", "year vs mileage")
head(data1_cor_BT)
quantile(data1_cor_BT[, "price vs year"], c(0.05,0.950)) ## CI for correlation between price and year
quantile(data1_cor_BT[, "price vs mileage"], c(0.05,0.950)) ## CI for correlation between price and mileage
quantile(data1_cor_BT[, "year vs mileage"], c(0.05,0.950)) ## CI for correlation between year and mileage
                                                            ## |abs| less than 0.8 -- no multicollinearity

cor.test(tc$Price, tc$Year,
         alternative = "two.sided",
         method = "pearson")

cor.test(tc$Price, tc$Mileage,
         alternative = "two.sided",
         method = "pearson")

cor.test(tc$Year, tc$Mileage,                 
         alternative = "two.sided",
         method = "pearson")
# bootstrap result:
#       5%       95% 
#0.4039420 0.4088868 
#        5%        95% 
#-0.4246398 -0.1296231 
#        5%        95% 
#-0.7677143 -0.2485699 


```

# display linearity -- yes, linear -- we can fit lm ?? not so sure
```{r}
ggplot(tc, aes(x = Year, y = Price)) + 
  geom_point() + 
  facet_wrap(~Category) 
plot(x = tc$Year, y = tc$Price)
plot(x = tc$Mileage, y = tc$Price)
```

# explore residual plot to see whether wild/residual bootstrap  -- increasing pattern between residual and year, decreasing pattern between residual and mileage -- so we use wild regression
```{r}
n = nrow(tc)
xbar = mean(tc$Price)
price = tc$Price
sd = sd(tc$Price)
res = (price - xbar)/(sd*sqrt(1-1/n))
plot(tc$Year, res)
plot(tc$Mileage, res)
```


# bootstrap regression for low end
```{r}
cate = "low-end"
temp = tc %>% 
  filter(Category == cate) %>% 
  select(Price, Year, Mileage)
fit = lm(Price~Year + Mileage, data=temp)
n1= nrow(temp)
y_predict = predict(fit)
B = 100
coeff_BT_wild = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict+rnorm(n1, sd=abs(fit$residuals))
  data1_BT = data.frame(years=temp$Year, mileage = temp$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild[,'ce mileage'],c(0.05,0.95))
plot(fit)
#100 bt
#      5%      95% 
#-1301810 -1247678 
#      5%      95% 
#629.0282 655.8358 
#         5%         95% 
#-0.05583089 -0.05299645 
```
# bootstrap regression for mid range
```{r}
cate2 = "mid-range"
temp2 = tc %>% 
  filter(Category == cate2) %>% 
  select(Price, Year, Mileage)
fit2 = lm(Price~Year + Mileage, data=temp2)
n2= nrow(temp2)
y_predict2 = predict(fit2)
B = 1000
coeff_BT_wild2 = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict2+rnorm(n2, sd=abs(fit2$residuals))
  data1_BT = data.frame(years=temp2$Year, mileage = temp2$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild2[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild2) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild2[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild2[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild2[,'ce mileage'],c(0.05,0.95))
#      5%      95% 
#-2537737 -2402189 
#      5%      95% 
#1207.071 1274.225 
#         5%         95% 
#-0.06585668 -0.05922462 

```
# bootstrap regression for high end 
```{r}
cate3 = "high-end"
temp3 = tc %>% 
  filter(Category == cate3) %>% 
  select(Price, Year, Mileage)
fit3 = lm(Price~Year + Mileage, data=temp3)
n3= nrow(temp3)
y_predict3 = predict(fit3)
B = 1000
coeff_BT_wild3 = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict3+rnorm(n3, sd=abs(fit3$residuals))
  data1_BT = data.frame(years=temp3$Year, mileage = temp3$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild3[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild3) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild3[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce mileage'],c(0.05,0.95))
#        5%       95% 
# -30520559 -23186964 
#      5%      95% 
# 11628.99 15268.70 
#        5%        95% 
#-1.5287048 -0.3142366      
```


# discard! brand: Ford, training 7/3

```{r}
tesla = tc %>% filter(Make == "Audi")
train_size = nrow(tesla) * 0.7  ## 246
set.seed(100)
trainingRows <- sample(1:nrow(tesla), 246)
train <- tesla[trainingRows, ]
test <- tesla[-trainingRows, ]
```

# discard! convert Mileage and Year to categorical 
```{r}
train$Mileage = factor(train$Mileage)
train$Year = factor(train$Year)
fit = lm(Price ~ Year + Mileage + City, data = train)
summary(fit)
```
# discard! convert make into factor -- one liner regression
```{r}
train_size = nrow(tc) * 0.7  ## 863129
set.seed(100)
trainingRows <- sample(1:nrow(tc), 863129)
train <- tc[trainingRows, ]
test <- tc[-trainingRows, ]

train$Make = factor(train$Make)
fit = lm(Price ~ Year + Mileage + Make, data = train)
summary(fit)
```

# kernel density 3d plot for high end car
```{r}
library(np)
library(rgl)
library(magick)
tc <- read_csv("new_tc.csv")
cate3 = "high-end"
temp3 = tc %>% 
  filter(Category == cate3) %>% 
  select(Price, Year, Mileage)
out <- capture.output(
  bw_temp3 <- npregbw(formula = Price ~ ordered(Year) + Mileage, data = temp3,
                         regtype = "ll", nmulti = 2)
)
summary(bw_temp3)
fit_temp3 <- np::npreg(bw_temp3)
summary(fit_temp3)
x_grid <- sort(unique(temp3$Year))
y_grid <- sort(unique(temp3$Mileage))
xy_grid <- expand.grid(x_grid, y_grid)
colnames(xy_grid) <- c("Year", "Mileage")
f = as.data.frame(xy_grid)
f_1 <- f %>% 
  group_by(Year) %>%
  arrange(Year)
fit_graph <- predict(fit_temp3, newdata = f_1)
f_year <- matrix(data=fit_graph, nrow=16, byrow=T)
open3d() 
plot3d(x = temp3$Year, y = temp3$Mileage, z = temp3$Price, xlab = "Year", ylab = "Mileage", zlab = "Price", col = "coral")
rgl::surface3d(x = x_grid, y = y_grid, z = f_year,
               col = "skyblue", alpha = 0.8, lit = FALSE)
rgl::rglwidget()
par3d(windowRect = c(20, 30, 500, 500))
movie3d(spin3d(), duration = 10, clean=FALSE,dir='/Users/claire/Desktop/test', type = "gif", convert=NULL)

```

# kernel density for mid range car
```{r}
cate2 = "mid-range"
temp2 = tc %>% 
  filter(Category == cate2) %>% 
  select(Price, Year, Mileage)

index = sample(nrow(temp2), 1000, replace = T)
sampletemp2 = temp2[index, ] 
out <- capture.output(
  bw_temp2 <- npregbw(formula = Price ~ ordered(Year) + Mileage, data = sampletemp2,
                         regtype = "ll", nmulti = 2)
)
summary(bw_temp2)
fit_temp2 <- np::npreg(bw_temp2)
summary(fit_temp2)
x_grid2 <- sort(unique(sampletemp2$Year))
y_grid2 <- sort(unique(sampletemp2$Mileage))
xy_grid2 <- expand.grid(x_grid2, y_grid2)
colnames(xy_grid2) <- c("Year", "Mileage")
f = as.data.frame(xy_grid2)
f_new <- f %>% 
  group_by(Year) %>%
  arrange(Year)
fit_graph <- predict(fit_temp2, newdata = f_new)
f_year <- matrix(data=fit_graph, nrow=20, byrow=T)
open3d() 
plot3d(x = sampletemp2$Year, y = sampletemp2$Mileage, z = sampletemp2$Price, xlab = "Year", ylab = "Mileage", zlab = "Price", col = "coral")
rgl::surface3d(x = x_grid2, y = y_grid2, z = f_year,
               col = "skyblue", alpha = 0.8, lit = FALSE)
rgl::rglwidget()
par3d(windowRect = c(20, 30, 500, 500))
movie3d(spin3d(), duration = 10, clean=TRUE,dir='/Users/claire/Desktop/test', type = "gif", convert=NULL)

```

# kernel density for lower car
```{r}
cate1 = "low-end"
temp1 = tc %>% 
  filter(Category == cate1) %>% 
  select(Price, Year, Mileage)
index1 = sample(nrow(temp1), 1000, replace = T)
temp1 = temp1[index1, ]
out <- capture.output(
  bw_temp1 <- npregbw(formula = Price ~ Year + Mileage, data = temp1,
                         regtype = "ll", nmulti = 2)
)
summary(bw_temp1)
fit_temp1 <- np::npreg(bw_temp1)
summary(fit_temp1)
x_grid1 <- sort(unique(temp1$Year))
y_grid1 <- sort(unique(temp1$Mileage))
xy_grid1 <- expand.grid(x_grid1, y_grid1)
colnames(xy_grid1) <- c("Year", "Mileage")
f = as.data.frame(xy_grid1)
f_new_1 <- f %>% 
  group_by(Year) %>%
  arrange(Year)
fit_graph <- predict(fit_temp1, newdata = f_new_1)
f_year <- matrix(data=fit_graph, nrow=20, byrow=T)
open3d() 
plot3d(x = temp1$Year, y = temp1$Mileage, z = temp1$Price, xlab = "Year", ylab = "Mileage", zlab = "Price", col = "coral")
rgl::surface3d(x = x_grid1, y = y_grid1, z = f_year,
               col = "skyblue", alpha = 0.8, lit = FALSE)
rgl::rglwidget()
par3d(windowRect = c(20, 30, 500, 500))
movie3d(spin3d(), duration = 10, clean=TRUE,dir='/Users/claire/Desktop/test', type = "gif", convert=NULL)
```



# linear low end
```{r}
fitlinearl1 = lm(Price ~ Year + Mileage, data = temp1)
x_grid_l1 <- unique(temp1$Year)
y_grid_l1 <- unique(temp1$Mileage)
xy_grid_l1 <- as.matrix(expand.grid(x_grid_l1, y_grid_l1))
colnames(xy_grid_l1) <- c("Year", "Mileage")
fl1 = as.data.frame(xy_grid_l1)
f_new_l1 <- fl1 %>% 
  group_by(Year) %>%
  arrange(Year)
fit_graphl1 <- predict(fitlinearl1, newdata = f_new_l1)
f_yearl1 <- matrix(data=fit_graphl1, nrow=22, byrow=T)
open3d()
plot3d(x = temp1$Year, y = temp1$Mileage, z = temp1$Price, xlab = "Year", ylab = "Mileage", zlab = "Price", col = "skyblue")
rgl::surface3d(x = x_grid_l1, y = y_grid_l1, z = f_yearl1, col = "pink", alpha = 0.8, lit = FALSE)
rgl::rglwidget()
par3d(windowRect = c(20, 30, 500, 500))
movie3d(spin3d(), duration = 10, clean=TRUE,dir='/Users/claire/Desktop/test', type = "gif", convert=NULL)
```

# linear mid range
```{r}
fitlinearl2 = lm(Price ~ Year + Mileage, data = temp2)
x_gridl2 <- unique(temp2$Year)
y_gridl2 <- unique(temp2$Mileage)
xy_gridl2 <- as.matrix(expand.grid(x_gridl2, y_gridl2))
fit_graphl2 <- predict(fitlinearl2, newdata = xy_gridl2)
plot3d(x = temp2$Year, y = temp2$Mileage, z = temp2$Price, xlab = "Year", ylab = "Mileage", zlab = "Price")
rgl::surface3d(x = x_gridl2, y = y_gridl2, z = matrix(fit_graphl2, nrow = 16, ncol = 711),
               col = "green", alpha = 0.25, lit = FALSE)
rgl::rglwidget()
```

# linear high end
```{r}
cate3 = "high-end"
temp3 = tc %>% 
  filter(Category == cate3) %>% 
  select(Price, Year, Mileage)
fitlinearl3 = lm(Price ~ Year + Mileage, data = temp3)
## ci bt
n3= nrow(temp3)
y_predict3 = predict(fitlinearl3)
B = 1000
coeff_BT_wild3 = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  y_bt = y_predict3+rnorm(n3, sd=abs(fitlinearl3$residuals))
  data1_BT = data.frame(years=temp3$Year, mileage = temp3$Mileage, price=y_bt)
  fit_BT = lm(price~years + mileage, data=data1_BT)
  coeff_BT_wild3[i_BT,] = fit_BT$coefficients
}
colnames(coeff_BT_wild3) = c("Intercept","ce year", "ce mileage")
quantile(coeff_BT_wild3[,'Intercept'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce year'],c(0.05,0.95))
quantile(coeff_BT_wild3[,'ce mileage'],c(0.05,0.95))
#        5%       95% 
# -22398988 -18434292 
#       5%       95% 
# 9270.439 11238.037 
#       5%       95% 
# -2.234915 -1.402934     
x_gridl3 <- sort(unique(temp3$Year))
y_gridl3 <- sort(unique(temp3$Mileage))
xy_gridl3 <- expand.grid(x_gridl3, y_gridl3)
colnames(xy_gridl3) <- c("Year", "Mileage")
fl3 = as.data.frame(xy_gridl3)
f_new_l3 <- fl3 %>% 
  group_by(Year) %>%
  arrange(Year)
lower <- function(x,y) {
  return (-22398988 + 9270.439*x - 2.234915*y)
}
upper <- function(x,y) {
  return (-18434292 + 11238.037 * x - 1.402934*y)
}
fit_upper <- matrix(data = upper(f_new_l3$Year, f_new_l3$Mileage), nrow = 16, byrow = T)
fit_lower <- matrix(data = lower(f_new_l3$Year, f_new_l3$Mileage), nrow = 16, byrow = T)
fit_graphl3 <- predict(fitlinearl3, newdata = f_new_l3)
f_yearl3 <- matrix(data=fit_graphl3, nrow=16, byrow=T)
open3d()
plot3d(x = temp3$Year, y = temp3$Mileage, z = temp3$Price, xlab = "Year", ylab = "Mileage", zlab = "Price", col = "coral")
rgl::surface3d(x = x_gridl3, y = y_gridl3, z = f_yearl3,
               col = "orchid", alpha = 0.25, lit = FALSE)
rgl::surface3d(x = x_gridl3, y = y_gridl3, z = fit_lower,
               col = "skyblue", alpha = 0.25, lit = FALSE)
rgl::surface3d(x = x_gridl3, y = y_gridl3, z = fit_upper,
               col = "skyblue", alpha = 0.25, lit = FALSE)
rgl::rglwidget()
```
```{r}
# empirical bootstrap
B = 10000
n = nrow(temp3)
coeff_emp_BT = matrix(NA, nrow=B, ncol=3)
for(i_BT in 1:B){
  w = sample(n,n,replace=T)
  data_emp_BT = temp3[w,]
  fit_BT = lm(Price ~ Year + Mileage, data=data_emp_BT)
  coeff_emp_BT[i_BT,] = fit_BT$coefficients
}
colnames(coeff_emp_BT) = c("Intercept","year", "mileage")
quantile(coeff_emp_BT[,'Intercept'],c(0.05,0.95))
quantile(coeff_emp_BT[,'year'],c(0.05,0.95))
quantile(coeff_emp_BT[,'mileage'],c(0.05,0.95))
```

